### 2020-02-10 paper/revisions/review-responses.tex

        \end{solution}

        %---------------------------------------------------------------------

        \question The authors did not conduct a sufficient literature survey on the fuel cycle transition studies for the SD-TMSR concept. Cui. et al., have made an extensive study on fuel cycle transition for SD-TMSR by employing LEU and Pu as the startup fissile materials and two transition modes (in-core transition and ex-core transition) (see Ref.: Transition toward thorium fuel cycle in a molten salt reactor by using plutonium, NUCL SCI TECH (2017) 28:152; Possible scenarios for the transition to thorium fuel cycle in molten salt reactor by using enriched uranium. Progress in Nuclear Energy 104 (2018) 75-84). The state "For the SD-TMSR concept, simulating the fissile core startup and fuel cycle transition with other fissile materials (except U-233) has not studied before" in page 3 is therefore inaccurate.
        \begin{solution}
                Thank you for your helpful review. The state "For the SD-TMSR concept, simulating the fissile core startup and fuel cycle transition with other fissile materials (except$^{233}$U) has not studied before" in page 3, has been removed.

                The following paragraphs have been added to the manuscript.\\
                Cui \emph{et al.} (2017-2018) analyzed the fuel transition from enriched $^{235}$U/Th and Pu/Th to $^{233}$U/Th for the MSBR-like design. Cui \emph{et al.} found that the fuel transition can be achieved by using: (1) enriched uranium with greater than 40\% enrichment, which rises the proliferation concerns; (2) Pu from LWR (burn-up of 60 GWd/t) spent nuclear fuel \cite{cui2017transition,cui2018possible}.

                The research effort described in \cite{cui2017transition} and
                \cite{cui2018possible} is most similar to the work
                presented in this paper. However, a few major differences
                are: (1) we used built-in truly continuous fuel depletion capabilities in
                SERPENT-2 \cite{aufiero2013extended} while Cui \emph{et al.} employed an in-house tool named MSR
                Reprocessing Sequence (MSR-RS) \cite{ZOU2015114} which couples with SCALE and employs a batch-wise approach (the burnup simulation stops at a given time and restarts with a new liquid fuel composition after removal of discarded materials and addition of fissile/fertile materials); (2) we investigated different initial loading cases (HALEU, Pu+HALEU, reactor-grade Pu, TRU and $^{233}$U) while Cui \emph{et al.} studied only uranium and Pu as the startup fissile materials; (3) we adopted the SD-TMSR core geometry optimized by Li \emph{et al.} \cite{li_optimization_2018} which is different from the MSBR-like model adopted by Cui \emph{et al.}; (4) Cui \emph{et al.} studied the high-enriched uranium (e$>$90) while we excluded it due to the unavailability of this material for use as an initial reactor fuel (non-proliferation issues) and because the performance of this material is expected to be similar to $^{233}$U; (5) Cui \emph{et al.} introduced two scenarios for thorium fuel cycle transition: a Breeding and Burning (B\&B) scenario and a Pre-breeding scenario while we adopting different approach by introducing two different feed mechanisms: thorium and non-thorium feed mechanism.

                In more detail, in current work, the radii of the fuel channels in the outer and inner zone are 5 and 3.5 cm, respectively, while in Cui \emph{et al.} work, all fuel channels have the same radius 2.03 cm, also, the side length of the graphite hexagonal prism is different in both models (7.5 instead of 10 cm in Cui model), total salt volume, density, and composition are different than those used in our study. In Cui's scenarios, the amount of Th is kept constant. $^{233}$U is fed only in (B\&B) scenario and extra fissile materials would have to be added into the core if the produced $^{233}$U is not enough to maintain the reactor criticality. In the Pre-breeding scenario, all the $^{233}$U produced from the decay of the extracted $^{233}$Pa stored outside the core until it reaches the required startup mass for a new reactor. Thus, the criticality of the core is maintained by refueling external fissile material. 
                In current work, using the thorium feed mechanism, we simultaneously feed Th and all or part of produced $^{233}$U from the \texttt{Pa-decay tank} into the core. The excess $^{233}$U (if exist) is stored outside the core. In the non-thorium feed mechanism, we continuously inject external heavy metals (HALEU, Pu mixed with HALEU, reactor-grade Pu, or TRU) and part of produced $^{233}$U from the \texttt{Pa-decay tank}. Here we feed a part of produced $^{233}$U because in some cases (e.g., TRU) refueling only external heavy metals failed to maintain the core criticality (limitation is dm$_{total fuel}$ $\leq$ 0.1\%). In the non-thorium feed mechanism, the amount of Th decreases during reactor operation while it is constant in the thorium feed mechanism.

### 2020-02-10 paper/revisions/review-responses.tex

        \question In the SD-TMSR burnup calculation, how are the depletion regions and burnup time step defined?
        \begin{solution}

        		  Thank you for the excellent point.

        		  About depletion regions definition: we used depletion mode 1 in SERPENT-2 (i.e. burn 1), this means that the fuel material is not divided and all occurrences of it are treated as a single depletion zone. The fuel is liquid -it well- mixed all the time due to turbulent flow and mixing in the pump.

        		  About burnup time step definition: it is time interval given in days. During burnup time step, the SD-TMSR core was maintained critical and total fuel mass was almost constant (dm $\leq$ 0.1\%). We started our calculations by 30 days' time step but we do not find significant changes in $k_{eff}$ when we increased the time step to 365 days. The maximum change in $k_{eff}$ was $\leq$ 0.02\%), therefore we adopted 365-day as a time step for all simulation in this work.

@@

        		  We changed the paragraph as follows:\\
        		  The fuel is liquid -it well- mixed all the time due to turbulent flow and mixing in the pump. We used depletion mode 1 in SERPENT-2 (i.e. burn 1), this means that the fuel material is not divided and all occurrences of it are treated as a single depletion zone.

        		  We changed the paragraph as follows:\\
        		  During the depletion step, the core was maintained critical and total fuel mass was almost constant (dm $\leq$ 0.1\%). To determine the appropriate depletion step size, we conducted a time step refinement study. We started with $\Delta t=30$days and gradually increased the depletion time step until the error in $k_{eff}$ became significant. For $\Delta t=1$year the maximum change in $k_{eff}$ was $\leq$ 0.02\%), therefore we adopted a constant, 365-day-long time step for all simulation in this work.

@@

        \end{solution}

        %--------------------------------------------------------------------

        \question Phenomenons presented in Figure 3 that K-eff for the case of Pu+HALEU firstly increases and then decreases, and the time to become subcritical for the case of reactor-grade Pu is longest (except the case of U-233) should be explained.
        \begin{solution}
                 Thank you for your insightful comment. Phenomenons presented in Figure 3 have been described in the manuscript.

                 The following paragraph has been added to the manuscript.\\
                 For Pu+HALEU case, $k_{eff}$ firstly increases during the first $\approx$ $3$ EFPY and then gradually decreases. This is due to the conversion of $^{238}$U from HALEU into fissile plutonium, which often produces the majority of the neutrons until the transition to Th/$^{233}$U fuel cycle. Figure~\ref{fig:diffPuAll811} shows the variation in fissile plutonium isotopes ($^{239}$Pu , $^{241}$Pu, and $^{243}$Pu) inventory during reactor operation. For TRU and reactor-grade Pu, the fissile plutonium isotopes are gradually depleted, during operating (the fissile material is dominated by Pu in these cases). For HALEU case, the fissile material is dominated by $^{235}$U, however, $^{238}$U is converted into fissile plutonium. Therefore, the mass of fissile plutonium increases during operating but its value still relatively low. For Pu+HALEU case, the mass of the fissile plutonium isotopes increases within the first $\approx$ $3$ EFPY and then decreases. The reactivity from the produced Pu is much greater than the depletion of the initial reactivity (due to the depletion of initial fissile isotopes), consequently, $k_{eff}$ increases during the first $\approx$ $3$ years and then decreases; the amount of $^{233}$U generated in the Pu+HALEU case is insufficient to maintain the reactor criticality and counteract neutron absorption in non-fissile isotopes.

                 The following paragraph has been added to the manuscript.\\
                 Figure~\ref{fig:keff1} shows that the decrease in $k_{eff}$ for reactor-grade Pu case occurs at the much later operating time than for other cases because the neutron spectrum in the reactor-grade Pu initial core is hardened; more $^{232}$Th is being converted to $^{233}$U. Nevertheless, this amount of the $^{233}$U is not enough to counteract the neutron absorption in the non-fissile Pu isotopes after much of the $^{239}$Pu and $^{241}$Pu is depleted.

@@

        \end{solution}

        %--------------------------------------------------------------------
        \question In page 19, why the Pa extraction time was selected to be 30s since Pa is neither the gaseous FPs nor the non-dissolved metals?
        \begin{solution}
                  Thank you for the question. Specifically, Pa is removed from the fuel by chemical reprocessing into the \texttt{Pa-decay tank} to decay and produce $^{233}$U. The target is to maintain the core criticality for a long period of time without any external feed of $^{233}$U. Figure~\ref{fig:Keff_U2332} shows how Reprocessing Period (RP) affects the $k_{eff}$ and amount of $^{233}$U in the \texttt{Pa-decay tank} for the TRU-started SD-TMSR. The fast extraction of Pa from the core (RP = $30$ $s$) leads to long operating time. Also, a relatively large amount of $^{233}$U can be obtained when we applied RP = $30$ $s$ against $10.59$ $d$. The RP equal to $10.59$ $d$ is not suitable for TRU and reactor-grade Pu cases because additional $^{233}$U was required to supply the core after startup. Therefore, we selected the RP = $30$ $s$. This produces $^{233}$U in the \texttt{Pa-decay tank} barely enough to operate the TRU-started SD-TMSR for a long period of time ($\approx$ 40 years). For reactor-grade Pu case, the produced $^{233}$U is sufficient to maintain the core criticality within the 60-years of operation (see Figure~\ref{fig:keff2}).

                  The following paragraph has been added to the manuscript.\\
                  The target is to maintain the core criticality for a long period of time without any external feed of $^{233}$U. Figure~\ref{fig:Keff_U2332} shows how Reprocessing Period (RP) affects the $k_{eff}$ and amount of $^{233}$U in the \texttt{Pa-decay tank} for the TRU-started SD-TMSR. The fast extraction of Pa from the core (RP = $30$ $s$) leads to long operating time. Also, a relatively large amount of $^{233}$U can be obtained when we applied RP = $30$ $s$ against $10.59$ $d$. The RP equal to $10.59$ $d$ is not suitable for TRU and reactor-grade Pu cases because additional $^{233}$U was required to supply the core after startup. Therefore, we selected the RP = $30$ $s$. This produces $^{233}$U in the \texttt{Pa-decay tank} barely enough to operate the TRU-started SD-TMSR for a long period of time ($\approx$ 40 years). For reactor-grade Pu case, the produced $^{233}$U is sufficient to maintain the core criticality within the 60-years of operation (see Figure~\ref{fig:keff2}).
        \end{solution}

### 2020-02-11 paper/results.tex

For Pu+HALEU case, K${_eff}$ firstly increases during the first $\approx$ $3$ EFPY and then gradually decreases. This is due to the conversion of $^{238}$U from HALEU into fissile plutonium, which often produces the majority of the neutrons until the transition to Th/$^{233}$U fuel cycle. Figure~\ref{fig:diffPuAll811} shows the variation in fissile plutonium isotopes ($^{239}$Pu , $^{241}$Pu, and $^{243}$Pu) inventory during reactor operation. For TRU and reactor-grade Pu, the fissile plutonium isotopes are gradually depleted, during operating (the fissile material is dominated by Pu in these cases). For HALEU case, the fissile material is dominated by $^{235}$U, however, $^{238}$U is converted into fissile plutonium. Therefore, the mass of fissile plutonium increases during operating but its value still relatively low. For Pu+HALEU case, the mass of the fissile plutonium isotopes increases within the first $\approx$ $3$ EFPY and then decreases. The reactivity from the produced Pu is much greater than the depletion of the initial reactivity (due to the depletion of initial fissile isotopes), consequently, K${_eff}$ increases during the first $\approx$ $3$ years and then decreases; the amount of $^{233}$U generated in the Pu+HALEU case is insufficient to maintain the reactor criticality and counteract neutron absorption in non-fissile isotopes.

Figure~\ref{fig:keff1} shows that the decrease in K${_eff}$ for reactor-grade Pu case occurs at the much later operating time than for other cases because the neutron spectrum in the reactor-grade Pu initial core is hardened; more $^{232}$Th is being converted to $^{233}$U. Nevertheless, this amount of the $^{233}$U is not enough to counteract the neutron absorption in the non-fissile Pu isotopes after much of the $^{239}$Pu and $^{241}$Pu is depleted \cite{betzler2016modeling}.

@@

Figures~\ref{fig:inventoryCCCC} and~\ref{fig:inventoryPu_TRUCCC} demonstrate 
the evolution of important nuclide inventories for the $^{233}$U, Pu, and TRU cases 
respectively. For the $^{233}$U case (Figure~\ref{fig:inventoryCCCC}), the mass of Pa in the fuel salt is almost constant and reaches 
$17.8$  $kg$ at the end of the operation time. 
Additionally, the mass of minor actinides (MA) and Pu increases with time. The level of Pu in the fuel salt correlates with the mass of the MA; however, MA need more time to reach equilibrium than Pu. Uranium inventory increases during 
operation and reaches equilibrium after $\approx$ $27$ years. Figure~\ref{fig:inventoryCCCC} shows that refueling the core with thorium helps 
maintain an almost constant inventory throughout the full operation time. 
For the Pu and TRU cases (Figure~\ref{fig:inventoryPu_TRUCCC}), the Pa extraction time was selected to be 30 s 
to avoid poisoning the core. Figure~\ref{fig:inventoryPu_TRUCCC} shows that the mass of Pa in the fuel salt 
is relatively low when compared to Pa mass in the $^{233}$U case (Figure~\ref{fig:inventoryCCCC}). Major 
elements for all three cases reach the equilibrium state after $\approx$ $30$ 
years (see Figure~\ref{fig:inventoryCCCC} and~\ref{fig:inventoryPu_TRUCCC}).
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{inventoryCCCC.png}
	\vspace{-0.4in}
	\caption{Evolution of important nuclide inventories for $^{233}$U 
		startup fuel (MA involves Np, Am, Cm) \cite{ashraf2019whole_core}.}
	\label{fig:inventoryCCCC}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{inventoryPu_TRUCCC.png}
	\vspace{-0.5in}
	\caption{Evolution of important nuclide inventories for the 
		reactor-grade Pu case (solid lines) and for the TRU case (dashed lines).}
	\label{fig:inventoryPu_TRUCCC}
\end{figure}

>>>>

For Pu+HALEU case, $k_{eff}$ firstly increases during the first $\approx$ $3$ EFPY and then gradually decreases. This is due to the conversion of $^{238}$U from HALEU into fissile plutonium, which often produces the majority of the neutrons until the transition to Th/$^{233}$U fuel cycle. Figure~\ref{fig:diffPuAll811} shows the variation in fissile plutonium isotopes ($^{239}$Pu , $^{241}$Pu, and $^{243}$Pu) inventory during reactor operation. For TRU and reactor-grade Pu, the fissile plutonium isotopes are gradually depleted, during operating (the fissile material is dominated by Pu in these cases). For HALEU case, the fissile material is dominated by $^{235}$U, however, $^{238}$U is converted into fissile plutonium. Therefore, the mass of fissile plutonium increases during operating but its value still relatively low. For Pu+HALEU case, the mass of the fissile plutonium isotopes increases within the first $\approx$ $3$ EFPY and then decreases. The reactivity from the produced Pu is much greater than the depletion of the initial reactivity (due to the depletion of initial fissile isotopes), consequently, $k_{eff}$ increases during the first $\approx$ $3$ years and then decreases; the amount of $^{233}$U generated in the Pu+HALEU case is insufficient to maintain the reactor criticality and counteract neutron absorption in non-fissile isotopes.

Figure~\ref{fig:keff1} shows that the decrease in $k_{eff}$ for reactor-grade Pu case occurs at the much later operating time than for other cases because the neutron spectrum in the reactor-grade Pu initial core is hardened; more $^{232}$Th is being converted to $^{233}$U. Nevertheless, this amount of the $^{233}$U is not enough to counteract the neutron absorption in the non-fissile Pu isotopes after much of the $^{239}$Pu and $^{241}$Pu is depleted \cite{betzler2016modeling}.

@@

Figures~\ref{fig:inventoryCCCC} and~\ref{fig:inventoryPu_TRUCCC} demonstrate 
the evolution of important nuclide inventories for the $^{233}$U, Pu, and TRU cases 
respectively. For the $^{233}$U case (Figure~\ref{fig:inventoryCCCC}), the mass of Pa in the fuel salt is almost constant and reaches 
$17.8$ $kg$ at the end of the operation time. 
Additionally, the mass of minor actinides (MA) and Pu increases with time. The level of Pu in the fuel salt correlates with the mass of the MA; however, MA need more time to reach equilibrium than Pu. Uranium inventory increases during 
operation and reaches equilibrium after $\approx$ $27$ years. Figure~\ref{fig:inventoryCCCC} shows that refueling the core with thorium helps 
maintain an almost constant inventory throughout the full operation time.

The target is to maintain the core criticality for a long period of time without any external feed of $^{233}$U. Figure~\ref{fig:Keff_U2332} shows how Reprocessing Period (RP) affects the $k_{eff}$ and amount of $^{233}$U in the \texttt{Pa-decay tank} for the TRU-started SD-TMSR. The fast extraction of Pa from the core (RP = $30$ $s$) leads to long operating time. Also, a relatively large amount of $^{233}$U can be obtained when we applied RP = $30$ $s$ against $10.59$ $d$. The RP equal to $10.59$ $d$ is not suitable for TRU and reactor-grade Pu cases because additional $^{233}$U was required to supply the core after startup. Therefore, we selected the RP = $30$ $s$. This produces $^{233}$U in the \texttt{Pa-decay tank} barely enough to operate the TRU-started SD-TMSR for a long period of time ($\approx$ $40$ years). For reactor-grade Pu case, the produced $^{233}$U is sufficient to maintain the core criticality within the 60-years of operation (see Figure~\ref{fig:keff2}).

Figure~\ref{fig:inventoryPu_TRUCCC} shows that the mass of Pa in the fuel salt 
is relatively low when compared to Pa mass in the $^{233}$U case (Figure~\ref{fig:inventoryCCCC}). Major 
elements for all three cases reach the equilibrium state after $\approx$ $30$ 
years (see Figure~\ref{fig:inventoryCCCC} and~\ref{fig:inventoryPu_TRUCCC}).

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{inventoryCCCC.png}
	\vspace{-0.4in}
	\caption{Evolution of important nuclide inventories for $^{233}$U 
		startup fuel (MA involves Np, Am, Cm) \cite{ashraf2019whole_core}.}
	\label{fig:inventoryCCCC}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{Keff_U2332.png}
	\vspace{-0.5in}
	\caption{The effect of Reprocessing Period (RP) on the $k_{eff}$ and amount of $^{233}$U in the \texttt{Pa-decay tank} for the TRU-started SD-TMSR.}
	\label{fig:Keff_U2332}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\textwidth]{inventoryPu_TRUCCC.png}
	\vspace{-0.5in}
	\caption{Evolution of important nuclide inventories for the 
		reactor-grade Pu case (solid lines) and for the TRU case (dashed lines).}
	\label{fig:inventoryPu_TRUCCC}
\end{figure}
